The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-24 paper/paper.Rnw
Added a bunch of posterior predictive checks for the models

<<default_conf>>=
library(knitr)
library(arm)
library(saves)
library(haven)
library(PISA2000lite)
library(PISA2003lite)
library(PISA2006lite)
library(PISA2009lite)
library(PISA2012lite)
library(intsvy)
library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
library(countrycode) # For region variable
library(car)
library(SAScii)
library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
library(lme4)
library(modelr)
library(ggrepel)
library(xtable)
library(ggthemes)
library(matrixStats)
library(WDI)
library(tidyverse)

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  mutate(num_tracks = factor(car::recode(tracks15y, "5 = 4")),
         age_selection = ifelse(selage >= 15, 1, 0))

library(rstanarm)

mod_tracking <- stan_glmer(
  difference ~
    num_tracks +
    age_selection + (1 | country),
  family = gaussian(),
  data = ready_data,
  prior = normal(0, 2),
  prior_intercept = normal(0, 5),
  prior_aux = cauchy(0, 1))

# mod_tracking <-
#   ready_data %>%
#   lmer(difference ~
#        num_tracks +
#        age_selection + (1 | country), data = .)

# new_data <-
#   data_grid(ready_data,
#             selage, num_tracks, central_examination, .model = mod_tracking) %>%
#   filter(complete.cases(.))

new_data <-
  ready_data %>%
  filter(!is.na(age_selection), !is.na(num_tracks)) %>%
  expand(country, age_selection, num_tracks) %>%
  filter(complete.cases(.))

@@

new_data$pred <- colMeans(predictions)
new_data$pred_low <- new_data$pred - (2 * sd_obs)
new_data$pred_high <- new_data$pred + (2 * sd_obs)

@@

>>>>

<<default_conf>>=
library(knitr)
library(arm)
library(saves)
library(haven)
library(PISA2000lite)
library(PISA2003lite)
library(PISA2006lite)
library(PISA2009lite)
library(PISA2012lite)
library(intsvy)
library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
library(countrycode) # For region variable
library(car)
library(SAScii)
library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
library(lme4)
library(modelr)
library(ggrepel)
library(xtable)
library(ggthemes)
library(matrixStats)
library(WDI)
library(rstanarm)
library(tidyverse)

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = ifelse(selage >= 15, 1, 0)) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

library(brms)

mod_tracking <-
  brms::brm(
  difference ~
    num_tracks +
    age_selection + (1 | country),
  family = gaussian(),
  data = ready_data,
  prior = c(set_prior("cauchy(2, 0.5)", coef = "num_tracks1"),
            set_prior("cauchy(13, 1)", coef = "age_selection")
            ),
  warmup = 1000, iter = 2000, chains = 5
  )

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model

new_data <-
  ready_data %>%
  filter(country %in% countries) %>%
  expand(country, age_selection, num_tracks) %>%
  filter(complete.cases(.))

@@

new_data <-
  new_data %>%
  mutate(
    pred = colMeans(predictions),
    pred_low = pred - (2 * sd_obs),
    pred_high = pred + (2 * sd_obs)
  )

@@

y_rep <- posterior_predict(mod_tracking)

replication <-
  map(1:2000, ~ {
  ready_data %>%
  mutate(pred_sample = y_rep[sample(1:4000, 1), , drop = TRUE]) %>%
  summarize(perc_sample = mean(pred_sample > difference),
            minus_sample = mean(pred_sample - difference))
})

replication %>%
  enframe() %>%
  unnest() %>%
  qplot(minus_sample, geom = "histogram", data = .)

replication %>%
  enframe() %>%
  unnest() %>%
  summarize(mean(between(minus_sample, -0.05, 0.05)))

pp_check(mod_tracking, nsamples = 100)

pp_check(mod_tracking, type = "error_scatter", nsamples = 100)
